---
title: "Tips to increase a profit"
author: "Dana Yerlanova"
date: "2019/04/05"
output: html_document
---

##By the way, I started to work with dataset where we should provide a tips to increase the profit of company.This dataset contains information of 2016-2018 sales. Our mission is to find what kind of factors affect to the profit and what company should do to acclerate its disposal.


###3.1 Loading libraries required and reading the data into R
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DT)
library(dplyr)
library(lubridate)
library(ggplot2)
library(reshape2)
library(lubridate)
library(rpart)
library(DMwR)
library(rpart.plot)
```

###Preview of dataset.
###Mutate new column Profit, which equals Fact..KZT..with.discount.- Cost.of.sales
```{r, echo = F}
df<-read.csv('C://Users/dyerl/Documents/table.csv', stringsAsFactors = F)
df$Date <- mdy(df$Date)
df$year <- year(df$Date)
df$month <- month(df$Date)
df$day <- day(df$Date)
df$weekday <- weekdays(df$Date)
options(scipen=100000)
dim(df)
str(df[,c(-10:-1)])
m<-sapply(df, function(x)sum(is.na(x)))
m
df$Cost.of.sales <- ifelse(is.na(df$Cost.of.sales), mean(df$Cost.of.sales, na.rm=TRUE), df$Cost.of.sales)
df<-df%>%mutate(Profit = Fact..KZT..with.discount.- Cost.of.sales)
tail(df$Cost.of.sales, 10)
```

###1.Plans of shop, and real profit in 2016, 2017, 2018
```{r}
df_year <- df%>% group_by(year)%>%summarize(meanProfit=mean(Profit), total=sum(Profit), planProfit= sum(Plan..KZT), difference = planProfit-total)
datatable(df_year)
options(scipen=100000)
mydata <- data.frame(Real_with_discount=df_year$total, Predicted=df_year$planProfit)
barplot(as.matrix(mydata), main="Real & Predicted income", ylab="Total", beside=TRUE, 
        col=terrain.colors(3))
legend("topright", c("2016", "2017","2018"), col=terrain.colors(3), lwd=15);
```

##2. Traffic
```{r}
df1<-df%>%group_by(year)%>%summarize(meanTraffic=mean(Traffic))
ggplot(data=df1, aes(year, meanTraffic))+geom_line()+geom_point()
```


#SEASON


##1.Let's check factor month, weekday. Maybe, having low income due to seasonality?
```{r}
df_sale_month <- df%>%group_by(month)%>%summarize(meanProfitMonth = mean(Profit),meanTrafficM = mean(Traffic),meanCustomerM = mean(Receipt..customers.))
df_sale_month
ggplot(data=df_sale_month, aes(factor(month), meanProfitMonth, fill=meanProfitMonth))+
  geom_bar(stat="identity")

```
###As a result, the most high selling month is October, the lowest is February

##So, let's try to look at profit for every season:
```{r}
df_sale_season <- df%>%mutate(Season = ifelse(month<=2|month==12, 'Winter', ifelse(month>=3&month<=5, 'Spring',ifelse(month>=6&month<=8, 'Summer','Fall'))))
df_season <- df_sale_season%>%group_by(Season)%>%summarize(meanP = mean(Profit),meanTr = mean(Traffic))
df_season
library(ggplot2)
ggplot(data=df_season, aes(Season,meanP))+
 geom_bar(stat="identity",fill = '#1E90FF', color ='#00008B')+
  ggtitle('Season histogram') +
  ylab('mean Profit for 2016-2017') + 
  xlab('Season')


```

###The highest profit in Fall, and high traffic during Spring.

##2. So, now let's check, how weekday affect to this measurements
```{r}
df_sale_max_days<-df%>%group_by(weekday)%>%summarize(meanProfit = mean(Profit), meanCustomer=mean(Traffic))
ggplot(data=df_sale_max_days, aes(weekday, meanProfit, fill=meanProfit, binwidth = 1))+
  geom_bar(stat="identity")
ggplot(data=df_sale_max_days, aes(weekday, meanCustomer, fill=meanCustomer, binwidth = 1))+
  geom_bar(stat="identity")

```

###Sunday, Saturday and Friday have the highest mean Profit and customers.

#So, we can do a conclusion, that we should do discounts, use ads during winter,spring to get more traffic.Also, make good deal during workdays, like 20% discount only on Thursdays.

#TRAFFIC

###Now, let's analyze does discount affect to traffic of customers or not:
```{r}
df_discount<-df%>%mutate(Discount_show=ifelse(Average.discount<30,'>>30%',ifelse(Average.discount>=30&Average.discount<=50,'30%-50%', '<<50%')))
df_discount_1<-df_discount%>%group_by(Discount_show)%>%summarize(AverageTraffic= mean(Traffic), Average_customer=mean(Receipt..customers.))
ggplot(data=df, aes(Average.discount,Traffic))+
  geom_point()+
  geom_smooth()
```
###We can see that discount from 30-50% works well, we should cut discount which over 60%


###Does high traffic mean that they become customers?
```{r}
ggplot(data=df, aes(Traffic, Receipt..customers.))+
  geom_point()+
  geom_smooth(model=lm)
```
### Yes, more traffic, more clients.

#We should make discount range from 30-50%. Also, should work with ads.

#DISCOUNT

## 6. How discount changes through month and weekday?
```{r}
df_disc_month<-df%>%group_by(month)%>%summarize(meanDiscount=mean(Average.discount), meanProfit= mean(Profit))
datatable(df_disc_month)
ggplot(data=df_disc_month, aes(month, meanDiscount, fill=meanProfit))+
  geom_bar(stat='identity')
df_disc_weekday<-df%>%group_by(weekday)%>%summarize(meanDiscount=mean(Average.discount), meanProfit= mean(Profit))
datatable(df_disc_weekday)
ggplot(data=df_disc_weekday, aes(weekday, meanDiscount, fill=meanProfit))+
  geom_bar(stat='identity')

```
## As a result, we can make a concclusion that the highest profit was in October, but it has the lowest discount average and ,medium amount of customers


##How discount affects to the profit of company
```{r}
summary(df$Diff..KZT)
summary(df$Average.discount)
ggplot(data= df, aes(Average.discount, Diff..KZT))+
  geom_point()+
  geom_smooth(method='loess')
```
###Basicly, we see that, discount doesn't affect to the Difference.Therefore, cutting discount doesn'twork

##Let's now check does having a high discount, leds to a big differeniation?
```{r}
options(scipen=100000)
ggplot(data=df_discount, aes(Profit, fill= Discount_show))+
  geom_histogram()+
  geom_vline(xintercept = 3)
```
### Discount didn't cost any changes, but itss good source to make high traffic

#AVERAGE PIECE PRICE

##How Average price piece with Difference
```{r}
ggplot(data=df%>%filter(Average.price.piece<40000),aes(Average.price.piece, Profit, color = Average.discount))+
  geom_point()+
   geom_smooth(method='loess')
ggplot(data=df%>%filter(Average.price.piece<40000),aes(Average.price.piece, Fact..KZT..with.discount., color = Average.discount))+
  geom_point()+
   geom_smooth(method='loess')
ggplot(data=df%>%filter(Average.price.piece<40000),aes(Average.price.piece, Fact..KZT..without.discount., color = Average.discount))+
  geom_point()+
   geom_smooth(method='loess')
```
##So we can made a conclusion that the profit increase, when Average piece price 0-10000, 15000-20000, profit increases.Therefore we should do discount to expensive items.

#PRODUCTIVITY

## How work worktime, productivity changes through a month? Does it affect to profit?
```{r}
df_day_month<-df%>%group_by(month)%>%summarize(meanTime=mean(Qty.of.hours),Profit=mean(Profit))
ggplot(data=df_day_month, aes(factor(month), meanTime, fill=Profit))+
  geom_bar(stat='identity')

df_day_month<-df%>%group_by(month)%>%summarize(meanProductivity=mean(Productivity),Cost.of.sales=mean(Cost.of.sales))
ggplot(data=df_day_month, aes(factor(month), meanProductivity, fill=Cost.of.sales))+
  geom_bar(stat='identity')

```

### How productivity, worktime changes through weekday? Does it affect to profit?
```{r}

df_day_weekday<-df%>%group_by(weekday)%>%summarize(meanTime=mean(Qty.of.hours),Profit=mean(Profit))
ggplot(data=df_day_weekday, aes(weekday,meanTime, fill=Profit))+
  geom_bar(stat='identity')

df_day_weekday<-df%>%group_by(weekday)%>%summarize(meanProductivity=mean(Productivity),Cost.of.sales=mean(Cost.of.sales))
ggplot(data=df_day_weekday, aes(weekday, meanProductivity, fill=Cost.of.sales))+
  geom_bar(stat='identity')

```
## So we can made a conclusion that productivity doesn't so much affect to the profit, but it rises cost of sales.Mean time helps to serve more customers.

## How discount affect to the other measures?
```{r}
library(rpart)
library(DMwR)
library(rpart.plot)
Discfactor <- lm(Average.discount ~  Qty.of.hours + Productivity + month+Average.price.piece+Traffic,
                     data = df)
Discfactor
```

###Let's see what kind of factors affect to the income.
```{r}
Factors <- lm(Profit ~ Average.discount + Qty.of.hours + Productivity + month+Average.price.piece+Traffic,
                     data = df)
Factors


```

##The hidhest profit month, might be it doesn't depend on month, month equal or not.
```{r}
library(lubridate)
t.test(df$Fact..KZT..with.discount.[df$month == 10 & df$year == 2016],
       df$Fact..KZT..with.discount.[df$month == 10 & df$year == 2017])

x<- df %>% select(Fact..KZT..with.discount., month, year)%>%filter(month==3, year == 2016|year==2017)
fligner.test(Fact..KZT..with.discount.~year, data=x)
```
#In conclusion:
#1. Do discounts when there isn't seasoning Winter-Spring
#2. Increasing traffic during Spring
#3. Increase quantity of hours of work during weekends
#4. Make average check 10000-15000
